---
title: "simulation_demo"
author: "Leanne Dong"
date: "13/07/2019"
output:
  pdf_document: default
  html_document: default
header-includes: \usepackage[ruled,vlined,linesnumbered]{algorithm2e}
                 \newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{blue}{#1}}
                 \SetCommentSty{mycommfont}
---

# Poisson process

We will simulate Poisson process by first principle. Algorithm taken from [epfl](http://transp-or.epfl.ch/courses/OptSim2012/slides/05b-poisson.pdf)

## Homogeneous Poisson Process

\begin{algorithm}[H]
\DontPrintSemicolon
\SetAlgoLined
\KwResult{Write here the result}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{ $T$ or $N$, $\lambda$}
%\Output{$S(t)$}
\textbf{Initialization} $T=0$, $k=0$, $S=vector()$.
\BlankLine
\While{ true\tcp*{$t<T$ or $n\le N$}
 }{   
    \textbf{Draw} $r\sim U(0,1)$.\;
    \textbf{Set} $t=t-\ln(r)/\lambda$.\;
    \eIf{$t_{n}>T$}{
    \text{return} $\{t_k\}_{k=1,2,\cdots,n-1}$\;
    }{
    $n=n+1$\;
    }
}
\caption{Simulation of event times of a Poisson process with rate $\lambda$ until time $T$
}
\end{algorithm} 

```{r}
rpoisson <- function(Tmax = NULL, Nmax = NULL, lambda) {
    
    # we can have both NULL or both set at the same time.
    if ( !xor(is.null(Tmax), is.null(Nmax)) ) stop("Need to set one (and only one) of Nmax or Tmax")
    
    t = 0
    k = 0
    S = vector()
    
    while (T) {
        r <- runif(1)
        t <- t - log(r) / lambda
        k <- k + 1
        S <- c(S, t)
        
        if (!is.null(Tmax) && (t >= Tmax)) break;
        if (!is.null(Nmax) && (length(S) >= Nmax)) break;
    }
    
    (S)
}
```

```{r}
#r1<-rpoisson(Tmax=10,lambda=1)
tmax=100
r1<-rpoisson(Tmax=tmax,lambda=1)
t_series <- seq(0,max(r1), length=length(r1))
#t_series <- seq(0,length(r1)-1,1)
plot(t_series,r1)
```

## Nonhomogeneous Poisson process


\begin{algorithm}[H]
\DontPrintSemicolon
\SetAlgoLined
\KwResult{Write here the result}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{ $T$ or $N$, $\lambda(t)$}
%\Output{$S(t)$}
\textbf{Initialization} $T=0$ or $n=m=0$, $\lambda^*=\sup_{0\le t\le T}\lambda(t)$.
\BlankLine
\While{ $\tau_m<T$ }
 {   
    \textbf{Draw} $r\sim U(0,1)$.\;
    \textbf{Set} $\tau_m=\tau_m-\ln(r)/\lambda^*$.\;
    \textbf{Draw} $s\sim U(0,1)$.\;
    \If{$s\le \frac{\lambda(\tau_m)}{\lambda^*}$}{
    \text{return} $t_n=\tau_m$\;
                  $n=n+1$\;
    }{
    $m=m+1$\;}
 }


\eIf{$t_n\le T$}{
\text{return} $\{t_k,k=1,2,\cdots,n-1\}$
}{
\text{return} $\{t_k,k=1,2,\cdots,n-2\}$
}


\caption{Simulation of event times of a Inhomogeneous Poisson process with rate $\lambda$ until time $T$
}
\end{algorithm} 

```{r}
rnhpoisson <- function(Tmax = NULL, Nmax = NULL, FUN, LambdaMax, ...) {
    
    # we can have both NULL or both set at the same time.
    if ( !xor(is.null(Tmax), is.null(Nmax)) ) stop("Need to set one (and only one) of Nmax or Tmax")
    if (LambdaMax == 0) return(NA) ## means we cannot get any event
    
    t = 0
    k = 0
    S = vector()
    
    while (T) {
        r <- runif(1)
        t <- t - log(r) / LambdaMax
        
        s <- runif(1)
        fnct <- FUN(t, ...)
        thr <- fnct/LambdaMax
        if ( s <= thr) {
            k <- k + 1
            S <- c(S, t)
        }
        
        ## exit loop at either time max, number max, or when the intensity at the current timestep is zero
        if (!is.null(Tmax) && (t >= Tmax)) break;
        if (!is.null(Nmax) && (length(S) >= Nmax)) break;
        # if (fnct < .Machine$double.eps) break;
        if (fnct < 10^-4) break; ## TODO: changed this so that simulation ends when no child
    }
    
    if (length(S) == 0) S <- NA
    return(S)
}
```


\begin{algorithm}[H]
\DontPrintSemicolon
\SetAlgoLined
\KwResult{Write here the result}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{ Tmax or Nmax, $\lambda(t)$}
\Output{$S(t)$}
\textbf{Initialization}  $t=0$, $k=0$, $S=vector()$.
\BlankLine
\While{ t<Tmax}{
    \textbf{Draw} $r\sim U(0,1)$.\;
    $t=t-\ln(r)/\lambda^*$.\;
    \textbf{Draw} $s\sim U(0,1)$. \textbf{Define} a new variable $\lambda(t)$ \;
    If $s\le \frac{\lambda(t)}{\lambda^*}$, set $k=k+1$ and $S(k)=t$\;
}
\caption{Simulation of event times of a Poisson process with rate $\lambda$ until time $T$
}
\end{algorithm} 







