---
title: "simulation_demo"
author: "Leanne Dong"
date: "13/07/2019"
output:
  pdf_document: default
  html_document: default
header-includes: \usepackage[ruled,vlined,linesnumbered]{algorithm2e}
---

# Poisson process

We will simulate Poisson process by first principle.

\begin{algorithm}[H]
\DontPrintSemicolon
\SetAlgoLined
\KwResult{Write here the result}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{ Tmax or Nmax, $\lambda$}
\Output{$S(t)$}
\textbf{Initialization} $t=0$, $k=0$, $S=0$.
\BlankLine
\While{ t<Tmax}{
    \textbf{Draw} $r\sim U(0,1)$.\;
    $t=t-\ln(r)/\lambda$.\;
    $k=k+1$, $S(k)=t$\;
}
\caption{Simulation of event times of a Poisson process with rate $\lambda$ until time $T$
}
\end{algorithm} 

```{r}
rpoisson <- function(Tmax = NULL, Nmax = NULL, lambda) {
    
    # we can have both NULL or both set at the same time.
    if ( !xor(is.null(Tmax), is.null(Nmax)) ) stop("Need to set one (and only one) of Nmax or Tmax")
    
    t = 0
    k = 0
    S = vector()
    
    while (T) {
        r <- runif(1)
        t <- t - log(r) / lambda
        k <- k + 1
        S <- c(S, t)
        
        if (!is.null(Tmax) && (t >= Tmax)) break;
        if (!is.null(Nmax) && (length(S) >= Nmax)) break;
    }
    
    (S)
}
```

```{r}
#r1<-rpoisson(Tmax=10,lambda=1)
tmax=100
r1<-rpoisson(Tmax=tmax,lambda=1)
#t_series <- seq(0,max(r1), by = max(r1)/tmax)
t_series <- seq(0,length(r1)-1,1)
plot(t_series,r1)
```








